---
title: "Week 3 Exercises"
author: "J.Blondeau"
date: "2023-09-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All the following question use the sttst_all_sites.xlsx file in the /data directory.

All answers should be made in the .Rmd file, showing the code and your answer and everything pushed back up to your GitHub repo.

triple stars renders a thin line across the page

***
1. Import sttstj_all_sites and inspect the data set.  What is the data structure and the data type of each field?


```{r}
 #this is a code chunk, anything inside the back tics is like being in the R console.

library(tidyverse)

d <- read_excel("data/sttstj_all_sites.xlsx", sheet=2)

#total number habitats
unique(d$hab)

#number pavement
PVMT<-d%>%filter(hab=="PVMT")

#number pavement or AGRF
PA<-d%>%filter(hab=="PVMT"|hab=="AGRF")

#number everything but pavement 
NoPVMT<-d%>%filter(hab!="PVMT")


#question 7

d7<-d%>%filter(hab=="AGRF"|hab=="SCR")%>%select(yr_site, hab)
head(d7)

#question 8
#arrange goes smallest to biggest, when do desc it goes biggest to smallest 
bdrk<-d%>%filter(hab=="BDRK")%>%arrange(desc(lat))
bdrk[1,]

#question 9
PA<-d%>%filter(hab=="AGRF"|hab=="PTRF")%>%arrange(lon)
PA[1,]

View(PA)

#question 10
split<-d %>% separate(yr_site, c('year', 'site'))
View(split)

#tell you the number of unique values 
length(unique(split$year))

#question 11

P04<-split%>%filter(year=="2004"& hab=="PVMT")%>%arrange(lat)
View(P04)
P04[1,]





```
***

2. How many samples are in the data set?  

2505 samples

3. How many unique habitat types are there?  What are the names of each habitat?

6, "PVMT" "AGRF" "SCR"  "PTRF" "BDRK" "HARD"

4. How many sites are there in the "PVMT" habitat type?

571 sites

5. How many sites are there in the "PVMT" or "AGRF" habitat types?

1538 sites

6. How many sites are there in all the habitat types EXCEPT "PVMT"?

1934 sites

7. Show the first 5 records (rows) of...
    - all sites sampled in "AGRF" or "SCR"
    - only show the "yr_site" and "hab" columns
    
    
      yr_site  hab  
  <chr>    <chr>
1 2001_243 AGRF 
2 2001_244 AGRF 
3 2001_247 AGRF 
4 2001_249 SCR  
5 2001_256 AGRF 
    
<!-- extra line break -->
<br>

8. What is the northern most site sampled in the "BDRK" habitat?
Northernmost is biggest number in latitude 
5450

9. What is the western most site sampled in either "AGRF" or "PTRF" habitat?

bigger negative number is more west 

9758

10. How many years were surveyed in the data set?

16

11. What is the southern most site sample in the "PVMT" habitat in 2004?

1932

